---
title: "2019 Canadian Federal Election Result if “everyone” had voted"
author: Ruoning Guo, Studnet No. 1004772114
date: "12/21/2020"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
mainfont: Times New Roman
fig.align: center
Url: https://github.com/ninaguoble/STA304_Final_Submission
---

```{r , echo=FALSE,results="hide",messange=FALSE,warnings=FALSE,include=FALSE}
library(readr)
library(dplyr)
library(plyr)
library(FrF2)
library(kableExtra)
library(ggplot2)
library(tidyverse)
library(broom) 
library(here)
library(skimr) 
library(brms)
library(ggpubr)
library(extrafont)
library(foreign)
library(srvyr)
poststrat <- read_csv("poststrat.csv")
survey <- read_csv("survey.csv")
head(survey)
kable(skim(survey))

library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```
## Abstract

MRP technique offers a robust alternative to traditional weighting techniques which corrects potential polling errors. With datasets from the Canadian Election Study Online Survey (2020), and Canadian General Social Survey (2017), I fit several logit models through multilevel regression and poststratiﬁcation (MRP) to investigate whether Andrew Scheer or Justin Trudeau would win the 2019 Canadian federal election if “everyone” had voted. Results show that the probability of Justin Trudeau to be elected is below 0.50. This might indicate that the CES might reflecting the general public’s voting intend which is not quite the same with the real election result.

## Keywords

Election poll forecast, Opinion surveys, Canadian Federal Election , Statistics Canada

## Introduction

One of the ongoing challenges in sampling election data and forecasting results is to better understand and reliably predict the diverse phenomena. Regardless of how elections are held, a proper voting process is crucial in determining whether a country's elected prime minister truly reflects the popular response. In many past elections, however, not everyone who was eligible to vote did so, which is a phenomenon that could affect the final outcome. In the 2016 United States presidential election, less than half of eligible Americans under 30 voted, which is far less than 71.4% of over-60s who voted (Macdonald, 2020). From report, only above three-quarters (77%) of Canadians voted in the 2019 federal election (Statistics Canada, 2020). Some doubt that Andrew Scheer might win if "everyone" voted (Brean, 2019).

It might be unrealistic for all Canadians to vote. Some votes would be more powerful than others, especially those cast in the most remote rural ridings (Lindsay J. M., 2020). Only Canadian citizens who are over 18 years old participated in the Campaign Period Survey (CPS) since not everyone would eligible to take part in the election (Election Canada, 2020). MRP has grown in popularity recently among scholars seeking to improve survey weighting techniques, particularly in forecasting state-level elections (Jonge et al., 2018). It has proven to be very powerful in many situations through partition the data into demographic cells and can be used to work out how many of each voter type there are in each constituency (Hutton, 2019). It estimates at the cell level using a multilevel regression model under the target population’s demographic composition (Downes & Carlin, 2020). In this report, I will use MRP to identify who would win the 2019 Canadian federal election if “everyone” had voted.

This study focuses on finding the predictability of election polls. Specifically, by using two datasets. One is the Campaign Period Survey (CPS) from Canadian Election Study (Stephenson et al., 2020) to build a model that predicts voting behavior. Then, Canadian General Social Survey (GSS) census data (Government of Canada, 2017) is used for post stratification. Detailed methods of investigation will be introduced in Methodology (Section 2), followed by Results (Section 3) including MRP analysis and outcomes. The study then expand in several parts: the rest of sections contain interpretation of results (Section 4), summary and conclusions, debates on potential weakness, and explorations about future of the investigation. The last section (Section 5) is references of all the works cited in this study.

## Methodology

### Data

In this section, we first discuss about the specific data needed for our methodology, then about the prediction model, and finally how the two datasets and the model are combined to create better prediction using MRP. 

The survey dataset, given the name $survey$, is generated by multiple steps of imputation to deal with inconsistent sample sizes, geographies and other biases from the original CPS survey data (2020). This raw data involves a biased, but large sample with 37,822 observations, from an opt-in online poll which was conducted during the preceding the 2019 prime minister election. All targets in this sample are stratified by four regions and balanced on gender and age within each region. Among them, 28% of respondents aged 1834, 33% aged 35-54 and 39% are 55 and higher. The data is initially aimed for a daily sample of 1000 respondents per day for the campaign period, whereas they increased targets during the last five days of the campaign in order to increase the total number of respondents (Stephenson et al., 2020). There are 21,762 observations after dropping NA and generating other variables.

The following list provides baseline characteristics of the survey data set, in order to compare similarity and characteristics with the poststratification dataset. Some variables are not used in the main investigation, i.e. MRP process but are important to be checked so as to find the most proper demographic cells. The variables in the survey dataset are:
1.	cps19_gender indicating gender of the targets in the survey data, including three categories: “A woman”, “A man” and “Other(e.g. Trans, non-binary, two-spirit, gender-queer)”.
2.	cps19_yob indicating targets’ year of birth, ranging from 1920 (99 years old in 2019) to 2001 (18 years old in 2019).
3.	cps19_province refers to the categorical variable of 10 provinces in Canada where the survey takers currently living.
4.	cps19_income_number, numerical variable of targets’ total household income before taxes for the year of 2018, which is measured in Canadian dollars.
5.	cps19_education contains 12 categorial variables of the highest level of education that survey takers have completed, ranging from “No schooling” to “Professional degree or doctorate” plus an option “Don't know/ Prefer not to answer”.
6.	cps10_votechoice, the choice of which party that the Canadian Citizens in the sample willing to vote for, including six parties: Liberal Party, Conservative Party, NDP, Bloc Québécois, Green Party, People's Party, plus two extra options: “Another party (please specify)”, and “Don't know/ Prefer not to answer”.  
7.	cps19_fed_id, the variable of 9 categories which indicates the self-identified membership of which Party in federal politics, including the six parties mentioned in 7. and three extra options: “Another party (please specify)”, “None of these” and “Don't know/ Prefer not to answer”.
8.	cps19_lead_int_113, the choice for which party leader(s) that the survey takers think is/are intelligent.
and two new variables generated from the original dataset:
9.	agesurvey, the age of survey takers, which is created by 2019 minus value of cps19_yob for each observation, so as to convert the year of birth to number of how many years they lived since birth.
10.	income_cat, grouping the survey takers by numeric incomes according to the variable cps19_income_number into 6 income levels: from “less than $25000” to “$125,000 and more”.
11.	edu, shortening the education levels of survey takers into three categories according to the variable cps19_education: “under bachelor”, “bachelor or college”, “above bachelor" and “other”.
12.	Trudeau, the binary variable which equals to 1 for people who think for Justin Trudeau is intelligent and 0 otherwise.
13.  age_group, where the variable indicating the which group the survey takers’ belong to among 4 categories, "ages60plus","ages45to59", "ages30to44", and "ages18to29".

The poststratification dataset, $poststrat$ is generated by GSS census data which is an annual survey designed to study the living conditions and well-being of Canadians (Government of Canada, 2017). The target population of this survey is all Canadians aged above 18; However, the sampling frame is Canadians older than 15 with functional phone number who lives in private households in all provinces excluding the 3 territories (Government of Canada, 2017). The sampling frame should be a good representation of the target population, therefore the $poststrat$ is expected to provide useful and reliable information about the living conditions of Canadians. The non-response questions have been named NA. This paper will randomly select 10,000 observations from the sampled population of 20,602 individuals, and conduct the study only using the 10,000 newly sampled observations.

Some baseline characteristics of the data set are listed below, which are: 
1.	sex, indicating gender of the samples in $poststrat$ dataset, including two categories: Male and Female.
2.	province, refers to the categorical variable of 10 provinces in Canada where the survey takers currently living.
3.	age, the age of each individuals in the sample, ranging from 18 to 80, after filtering those who are under 18 years old.
4.	income_family, categorial variable which groups the samples into 6 income levels: from “less than $25000” to “$125,000 and more”.
5.	education,categorial variable which groups the samples into 5 education levels.
6.	p_edu, education levels of individuals into three categories: “under bachelor”, “bachelor or college”, “NA" and “other”.
7. age_group, new variable generated by the variable “age”, with 4 categories that are the same with which of survey data.

The following part is the comparison of demographics of the survey data and poststratification data, also the creation of new variables is processed after the comparision.

#### Figure 0a: histogram for sex variables of survey data and poststrat data

```{r,fig.align="center"}

surveysex<-ggplot(survey, aes(cps19_gender)) +
  geom_bar(fill = "#0073C2FF") +
  theme(text = element_text(size = 8))   
postsex<-ggplot(poststrat, aes(sex)) +
  geom_bar(fill = "#0073C2FF") +
  theme(text = element_text(size = 8))   
figure_sex <- ggarrange(surveysex, postsex,
                    labels = c("SurveyData", "PostStratData"),
                    ncol = 2, nrow = 1)
figure_sex
```

#### Figure 0b: histogram and boxplots for distribution of age variables of survey data and poststrat data

```{r}
#age
survey <- survey %>% mutate(agesurvey = 2019-survey$cps19_yob)
agesurvey <- 2019-survey$cps19_yob 
agepost <- poststrat$age
par(mfrow=c(2,2))
hist(agesurvey)
boxplot(agesurvey, horizontal=TRUE)
hist(agepost)
boxplot(agepost, horizontal=TRUE)
```

#### Figure 0c: The frequency barplots of province variables of survey data and poststrat data
```{r, fig.align="center"}
#province
surveyregion<-ggplot(survey, aes(cps19_province )) +
  geom_bar(fill = "#0073C2FF") 
postregion<-ggplot(poststrat, aes(province)) +
  geom_bar(fill = "#0073C2FF")
figure_province <- ggarrange(surveyregion, postregion,
                    labels = c("SurveyData", "PostStratData"),
                    ncol = 2, nrow = 1)
figure_province
```

#### Figure 0d: barplots of income categorial variable of survey data and poststrat data


```{r,fig.align="center", warning=FALSE }
#incomes-need to clean into uniformed categories
survey <- survey %>% mutate(income_cat = case_when(cps19_income_number  >= 125000 ~ "125k+",
cps19_income_number  >= 75000 & cps19_income_number  <= 124999 ~ "75-125k",
cps19_income_number  >= 50000 & cps19_income_number  <= 74999 ~ "50-75k",
cps19_income_number  >= 25000 & cps19_income_number  <= 49999 ~ "25-50k",
cps19_income_number < 25000 ~ "below25k"))

surveyinc<-ggplot(survey, aes(income_cat)) +
  geom_bar(fill = "#0073C2FF") 

p_income_cat <- c("Less than $25,000"="below25k","$25,000 to $49,999"="25-50k","$50,000 to $74,999"="50-75k","$75,000 to $99,999"="75-125k","$125,000 and more"="125k+","$100,000 to $ 124,999"="75-125k")
poststrat$income_family <-  revalue(poststrat$income_family, p_income_cat)

postinc<-ggplot(poststrat, aes(income_family)) +
  geom_bar(fill = "#0073C2FF")
figure_income <- ggarrange(surveyinc, postinc,
                    labels = c("SurveyData", "PostStratData"),
                    ncol = 2, nrow = 1)
figure_income
```


#### Figure 0e: histogram for different categories of education levels of survey data and poststrat data


```{r, fig.align="center",warning=FALSE}
#education
survey <- survey %>% mutate(edu = case_when(cps19_education  ==  "No schooling"  ~ "under bachelor",  cps19_education =="Some elementary school " ~ "under bachelor", 
cps19_education ==  "Completed elementary school" ~ "under bachelor", 
cps19_education ==" Some technical, community college, CEGEP, College Classique " ~ "	bachelor or college", 
cps19_education == " Completed technical, community college, CEGEP, College Classique "~ "	bachelor or college", 
cps19_education == "Some university"  ~ "	bachelor or college", 
cps19_education == "Bachelor's degree "~ "	bachelor or college", 
cps19_education ==  "Master's degree " ~ "above bachelor", 
cps19_education == "Professional degree or doctorate"~ "above bachelor",
cps19_education == "Don't know/ Prefer not to answer "~"other"))

poststrat <- poststrat%>% mutate(p_edu = case_when(education == "High school diploma or a high school equivalency certificate" ~ "under bachelor",
education == "Less than high school diploma or its equivalent"~"under bachelor",
education == "University certificate or diploma below the bachelor's level"~"under bachelor",
education == "Bachelor's degree (e.g. B.A., B.Sc., LL.B.)"~"bachelor or college",
education == "Trade certificate or diploma"~"other",
education == "College, CEGEP or other non-university certificate or diploma"~"bachelor or college",
))

surveyedu<-ggplot(survey, aes(edu)) +
  geom_bar(fill = "#0073C2FF") 
postedu<-ggplot(poststrat, aes(p_edu)) +
  geom_bar(fill = "#0073C2FF")
figure_edu <- ggarrange(surveyedu, postedu,
                    labels = c("SurveyData", "PostStratData"),
                    ncol = 1, nrow = 2)
figure_edu

```


#### Figure 0f: Compare survey data set's Party vote choice

```{r, fig.align="center",warning=FALSE}

survey$cps19_votechoice<- iconv(survey$cps19_votechoice, "ASCII", "UTF-8", sub="byte")
ggplot(survey, aes(x = factor(cps19_votechoice))) +
    geom_bar()
```

#### Figure 0g: Compare survey data set's Federal Party ID
```{r,  fig.align="center", warning=FALSE}

survey$cps19_fed_id<- iconv(survey$cps19_fed_id, "ASCII", "UTF-8", sub="byte")
ggplot(survey, aes(x = factor(cps19_fed_id))) +
    geom_bar()

```

#### Table 1 : Summary of the cleaned two datasets: Poststrat and Survey Data

```{r}
head(poststrat)
head(survey)
```
### Model

Since we are interested in whether or not the voters prefer to vote for Justin Trudeau, $Trudeau$ can be considered as a response variable for these data and a logistic regression analysis could be carried out to determine the effect of cps19_gender and age_group expected on the odds of voting for Trudeau. Under classic logistic regression, the estimator of $\beta_1$ would have a fixed value regardless of the value of predictors. However, in reality, people with different gender might have different $\beta_1$. 

Firstly, we are running an OLS regression with Trudeau as the dependent variable and cps19_gender and age_group as the independent variables. The dataset that we are using is survey. We are then saving that OLS regression to a variable called 'model'. We just used regular OLS even though our dependent variable is a binary. 

#### Table 2: OLS linear model summary

```{r}
model <- lm(Trudeau ~ cps19_gender + age_group, 
            data = survey)
broom::tidy(model) 
```
Now we would like to see if we can use what we found in the poll to get an estimate for each state based on their demographic features. This problem could be solved using the Bayesian approach, by incorporating our initial beliefs into the model (Kana, 2020). Therefore, Bayesian logistic regression model is the most appropriate model for this analysis. In this example, the first stage of the MRP modeling process implements a pair of multilevel logistic regression models predicting being a likely voter and supporting an individual candidate. For each gender and age_group we want the relevant coefficient in the poststrat and we can construct the estimates.

The logistic regressions are fitted with age_group and cps19_gender as predictor variables, where:
- cps19_gender: the variable indicating the gender of survey takers, which has two categories, Male and Female.
- age_group: the grouped categorical variable which separates individuals in the sample into four age groups: 18-29, 30-44, 45-59 and 60+.

```{r, include=FALSE}
library(dplyr)
library(plyr)
library(srvyr)
data(survey, package="survey")
# compute for each cell
survey1 <- survey %>% group_by(age_group, income_cat)

# compute population in 5 income status
library(data.table)
#as.data.table(poststrat1)[, sum(incomef), by = income_family]

incomef<- survey1$income_cat
incomef<- as.data.frame(incomef)
#as.data.table(incomef)[,sum(n),by=incomef]
incomef$incomef[incomef$incomef%in% "below25k"] <- "2198"
incomef$incomef[incomef$incomef %in% "25-50k"] <- "3959"
incomef$incomef[incomef$incomef%in% "50-75k"] <- "4749"
incomef$incomef[incomef$incomef %in% "75-125k"] <- "6620"
incomef$incomef[incomef$incomef %in% "125k+"] <- "4236"

survey1 = cbind(survey1,incomef)
survey1<-survey1%>%mutate(estimate = as.numeric(as.character(incomef))/21762)
               
```
Here we are making predictions using our model with some new data from the survey, and we saving the results of those predictions by adding a new column to the survey dataset called 'estimate'.
```{r}
survey1$estimate1 <-  model %>% predict(newdata = survey1)
```


```{r,include=FALSE}
survey2<- survey1%>% 
  select(income_cat,age_group,estimate,estimate1,edu)%>%
  mutate(Trudeau_predict_prop = estimate1*estimate)
```

The Baysian logistic model is: 

$$ log(p_i/(1-p_i)) = \beta_0+\beta_1  X_{cps19gender,i} + \beta_2  X_{18to29,i} +\beta_3  X_{30to44,i} +\beta_4  X_{45to59,i} + \beta_5  X_{60+,i}+\epsilon$$
Where:

- $X_{cps19gender}$ = 1 if the ith voter is a woman, and 0 if otherwise;
- $X_{18to29}$ = 1 if the i-th voter's age between 18 to 29 years old, and 0 if not 18 to 29 years old;
- $X_{30to44}$ = 1 if the i-th voter's age between 30 to 44 years old, and 0 if not 30 to 44 years old;
- $X_{45to59}$ = 1 if the i-th voter's age between 45 to 59 years old, and 0 if not 45 to 59 years old;
- $X_{60+}$ = 1 if the i-th voter's age above 60 years old, and 0 if not above 60 years old;
- $p_i$ is the probability of having a preference for voting Justin Trudeau.
- Coefficients $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$, $\beta_5$ represent change in log odds.For example, $\beta_1$ coefficient represents change in log odds for every one unit increase in $X_{cps19gender}$.


```{r}
modelbrm <- brm(Trudeau~ cps19_gender + age_group, 
             data = survey1, 
             family = bernoulli(),
             file = "C:/Users/Alienware/Desktop/STA304 Final"
             )
saveRDS(modelbrm, "C:/Users/Alienware/Desktop/STA304 Final/brms_model.rds")

```



#### Table 3: Model summary

```{r,result=TRUE,message=FALSE,warning=FALSE}

summary(modelbrm)
```

The output shows above is our partial regression output from the Bayesian model, where R uses "genderAwoman" and "age 18to29" as the level of the reference. Now we have the "estimates" values which are equivalent to beta coefficients in the logistic formula. A rough probability of voting for Justin Trudeau can be calculated based on the voter's cps19_gender and age_group.


## Results

#### Figure 1: plot model convergence to check whether there is evidence of non-convergence for the four chains.
```{r, fig.align="center",echo=FALSE,message=FALSE,warning=FALSE}
stanplot(modelbrm, 
         type = "trace")
```
To check non-convergence, we use the stanplot function from the brms package to plot the caterpillar plot for each parameter of interest. From the series of plot above, we find the four chains mix well (Guo et al., 2020).

#### Figure 2: check autocorrelation due to the fact that the presence of strong autocorrelation would bias variance estimates.
```{r, echo=FALSE,message=FALSE,warning=FALSE}
stanplot(modelbrm,type = "acf_bar")
```

The plots above show that the autocorrelation parameters all diminish to around zero (Guo et al., 2020).

#### Figure 3: visualise the point estimates and their associated 95%-CI
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.align="center"}
stanplot(modelbrm, 
         type = "areas",
         prob = 0.95)
```
In the Bayesian model, the 95% confidence interval states that there is 95% chance that the true value falls within this interval. Since the 95%-CI does not contain zero (except the second wor that contains the feature cps19gender_Awoman), the respective model parameters are likely meaningful.

#### An estimate for each employment status based on their demographic features.
```{r,result=TRUE,message=FALSE,warning=FALSE}
survey3 <- survey1%>% select(cps19_province, estimate1)%>% group_by(cps19_province)
x <- survey3
x %>%
  group_by(cps19_province) %>%
  dplyr::summarise(mean = mean(estimate1))
```

We estimate that the proportion of voters whose reside in Nunavut in favor of voting for Trudeau to be 0.447, which is the lowest; While voters in favor of voting for Trudeau to be 0.467 in Prince Edward Island, which is the highest. However, which province have which the difference in mean of voting probability is not significant. This is based off our post-stratification analysis of the proportion of voters in favour of voting for Justin Trudeau modeled by a Bayesian logistic regression model, which accounted for age_group and cps19_gender.

### Table 4: Estimates in terms of income status

```{r,result=TRUE,message=FALSE,warning=FALSE}

survey4 <- survey1%>% select(cps19_province, estimate1)%>% group_by(income_cat)
y <- survey4
y %>%
  group_by(income_cat) %>%
  dplyr::summarise(mean = mean(estimate1))

```

We estimate that the proportion of voters whose family earning at 25-50k dollars in 2018 (the second least tier) in favor of voting for Trudeau to be 0.468, which is the highest; While the richest voters (those family income at 125k+ in 2018) in favor of voting for Trudeau which is 0.451; This is based off our post-stratification analysis of the proportion of voters in favour of voting for Trudeau modeled by a Bayesian logistic regression model, which accounted for age_group and cps19_gender.

#### An estimate for each age groups based on their demographic features.
```{r,result=TRUE,message=FALSE,warning=FALSE}
survey5 <- survey1%>% select(age_group, estimate1)%>% group_by(age_group)
x1 <- survey5
x1 %>%
  group_by(age_group) %>%
  dplyr::summarise(mean = mean(estimate1))
```
We estimate that the proportion of voters whose ages between 19 and 29 in favor of voting for Trudeau to be 0.4656, while voters in favor of voting for Trudeau to be 0.4052 who aged between 30 to 44, which is the lowest probability among the four groups. However, the highest probability of voting to Trudeau can be seen among those who aged above 60 years old, which is 0.5138. This is based off our post-stratification analysis of the proportion of voters in favour of voting for Justin Trudeau modeled by a Bayesian logistic regression model, which accounted for age_group and cps19_gender.

We got the $\hat{y}$, now calculate for $\hat{y}^{PS}$, which is 0.588256.
```{r,echo=TRUE, result=TRUE,message=FALSE,warning=FALSE}
#Number of "125k+": 4236
#Number of "75-125k": 6620
#Number of "50-75k": 4749
#Number of "25-50k":3959
#Number of "below25k":2198

y_hat_ps=(0.4516233*4236+0.4562982*6620+0.4627370*4749+ 0.4683544*3959 + 0.4651849*2198)/(4236 + 3959 + 6620 +2198)
y_hat_ps

```

## Discussion

### Summary

#### Checking the evidence of non-convergence and autocorrelation for the four chains(age_group).

- Figure 1: there seems no evidence of non-convergence since the four chains mix well.

- Figure 2: As the autocorrelation parameters all diminish to around zero, there is no evidence of autocorrelation for all model variables in our four chains. 

- Figure 3: Gender and age group are weak but still meaningful predictors as for all graphs (maybe except the one which indicates genderAwoman), their 95% confidence intervals do not contain zero. In addition, gender is a more credible predictor in the constructed model since its density plot for woman is relatively narrower, while another graph indicating other genders except woman is wide. Thus, we can safely illustrate the model output. 

#### Model summary and Relating model with the real-world

In survey data, the population of male voters and voters having other genders is almost consistent with female voters(respectively 9814 males, 136 other and 11312 women), which also matches with the current population structure in Canada (Statistics Canada, 2020). However, Since CIs for all Population-Levels do contain zero(means p-value>0.05), there is no statistical evidence to show that preference for voting Justin Trudeau is associated with both cps19_gender and age_group of the voters. While as the age group increases, the coefficient tends to increase on average. In other words, if holding everything else constant, turnout voters who belong to the higher age group are more likely to vote for Trudeau. 

Nevertheless, this result is not reliable due to the bias of selection of the population itself. The number of senior voters turnout in 2019 federal election would be higher than that of those younger voters (Elections Canada, 2020). This may influence the result that there is a naturally higher participation rate in the federal election for those who are relatively older. In fact, voters aged 65 to 74 overall had the highest participation rate of fedral election in 2019, with 79.1 per cent casting a ballot in 2019, up from 78.8 per cent in 2015 (Elections Canada, 2020).

Probabilistic estimates for each strat are 0.4516233(125k+),	0.4683544(25-50k), 0.4627370(50-75k) 0.4562982(75-125k), and 0.4651849(below25k) for the five income categories (Table 4). It shows slight differences between strats while people not having very high family income are more likely to vote for Trudeau as re-elected president than those who are with higher family income or those who are the poorest. One thing that might explain this result is the policy that Liberal Party (which Justin Trudeau leads) claimed to the public. For example, while Andrew Scheer’s tax plan would give more to wealthy Canadians, the plan of Liberal Party lowers taxes the most for people who make less, helps the middle class, and ensures that the wealthy don’t get an extra hand up. The tax cut would also help lift twice as many Canadians out of poverty than the Conservative plan would (Liberal Party of Canada). Some financial policies made by Liberal are persistently favored by general publics. During the pandamic, a guaranteed basic income for all Canadians has emerged as the top policy choice of Liberal MPs (CBC, 2020). Especially for the people who are in the middle class and those who have lower family incomes, the Liberals are proposing a tax cut, where a middle class family gets 600 dollars and the wealthiest one per cent get zero. It will make sure the middle class pay no federal taxes on the first 15,000 dollars they earn, by raising the Basic Personal Amount (BPA) by almost 2,000 dollars for people earning under 147,000 dollars a year. It is designed to put millions of middle-class families on the path to a secure retirement(Liberal Party of Canada). 

On the contrary, the ratios are not differ very much in each tier of incomes. Some claims that the Liberal plan will do more to lift low-income Canadians out of poverty, according to analysis by B.C. economist Kevin Milligan, but for median-income households, the impact will be roughly the same (Urback, 2019).

Eventually, we forecast that Trudeau has a 58.8% chance to win the 2019 Canadian federal election based off our post-stratification estimates weighted by the three strats(calculation can be found at the end of the appendix). 

### Weaknesses & Next Steps

Since our research model contains only four variables, including two variables (gender and age groups) in the primary model and two variables(province and income status) in the stratification section. One drawback might be the lack of cells since there are only 25 cells. We probably need to add more categories or variables to see whether the brm result may change or not. 

Besides, the allocation of sample units to strata cannot be controlled since the variation of sample sizes according to strata could cause too few samples to categorize a reliable stratum mean and standard error (Westfall, Patterson& Coulston, 2011). 

Another drawback of a lack of variables in the model is of the problem: Omitted Variable Bias(OVB) which is a bias that stems from the absence of relevant variables in a model (S, 2020). To solve this, we need to add more variables and test each of their robustness to eliminate an underfit model. We might also need to compare with the actual election results and do a post-hoc analysis (or at least a survey) of how to better improve estimation in future elections, because of the nature of the large uncertainty of our post stratified estimates. 

In addition, the drawback of surveys and census themselves may exist. Since the model cannot reject the null hypothesis, the selection of variables is also need to be reconsidered. Perhaps a longer model containing a wider range of demographic characteristics is necessary. For example, adding educational levels and province variables in the brm model.

The next steps of this research may contain: 

- adding layers to do a Multilevel Regression Poststratification, therefore to achieve the adjustment for selection bias (Guo et al., 2020)
- correcting for imbalances in sample composition, even when the dataset involves a large number of variables--together with multilevel regression (Si, 2020).

## References

Alexander, R. (2019, December 03). Getting started with MRP. Retrieved from https://rohanalexander.com/posts/2019-12-04-getting_started_with_mrp/

Brean, J. (2019, October 21). How the 2019 federal election became a vote for nothing. nationalpost. https://nationalpost.com/feature/federal-election-2019-liberals-conservatives-ndp-green-ppc-bq.

CBC/Radio Canada. (2020, September 12). Guaranteed basic income tops policy priorities for Liberal caucus at upcoming convention | CBC News. CBCnews. https://www.cbc.ca/news/politics/guaranteed-basic-income-priorities-liberals-1.5721943.

Downes, M., &amp; Carlin, J. B. (2020). Multilevel Regression and Poststratification Versus Survey Sample Weighting for Estimating Population Quantities in Large Population Health Studies. American Journal of Epidemiology, 189(7), 717–725. https://doi.org/10.1093/aje/kwaa053 

Elections Canada. (2020, August 6). Voter Turnout by Sex and Age. https://www.elections.ca/content.aspx?section=res&amp;dir=rec/eval/pes2019/vtsa&amp;document=index&amp;lang=e#ftn1.

Election Canada. (2020, December 03). Who is eligible to vote in the federal election and referendums? FAQs on Voting. https://www.elections.ca/content.aspx?section=vot.

Government of Canada, S. (2017, February 27). The General Social Survey: An Overview. Retrieved October 19, 2020, https://www150.statcan.gc.ca/n1/pub/89f0115x/89f0115x2013001-eng.html

Guo, R., Zhu, Y., Guo, K., &amp; Meng, F. Bayesian logistic regression and post-stratification for estimating the outcome of 2020 Presidential Election: Will Donald Trump win? GitHub. https://github.com/YvonneYifanZhu/STA304_PS3_Submission.

Liberal Party of Canada. More Money, Tax Free: Our Platform. Go to Liberal Party of Canada. https://liberal.ca/our-platform/more-money-tax-free/.

Lindsay J. M. (2020, December 15). The 2020 Election by the Numbers. Council on Foreign Relations. Council on Foreign Relations. https://www.cfr.org/blog/2020-election-numbers.

Hutton, R. (2019, November 27). What Is the ‘MRP Poll’ and Can It Predict the U.K. Election? Bloomberg.com. https://www.bloomberg.com/news/articles/2019-11-27/what-is-the-mrp-poll-and-can-it-predict-the-u-k-election.

Jonge, C. P. K. D., Langer, G., &amp; Sinozich, S. (2018). Predicting State Presidential Election Results Using National Tracking Polls and Multilevel Regression with Poststratification (MRP). Public Opinion Quarterly, 82(3), 419–446. https://doi.org/10.1093/poq/nfy023

Kana, M. (2020, February 21). Introduction to Bayesian Logistic Regression. Retrieved from https://towardsdatascience.com/introduction-to-bayesian-logistic-regression-7e39a0bae691

Macdonald, B. (2020, November 02). Less than half of young Americans voted in the last presidential election. Our World in Data. https://ourworldindata.org/usa-electoral-turnout.

Rohan Alexander. (2020, February 11). A review of 'Forecasting elections with non-representative polls'. Rohan Alexander. https://rohanalexander.com/posts/2020-02-11-a-review-of-forecasting-elections-with-non-representative-polls/.

Statistics Canada. (2020, February 26). Reasons for not voting in the federal election, October 21, 2019. The Daily&nbsp;-&nbsp;. https://www150.statcan.gc.ca/n1/daily-quotidien/200226/dq200226b-eng.htm.

Si, Y. (2020). Multilevel Regression and Poststratification. Retrieved from http://www-personal.umich.edu/~yajuan/files/MrPbrownbag-YAJUANSI.pdf

Statistics Canada. (2020, December 20). Population estimates, quarterly. Statistics Canada: Canada's national statistical agency / Statistique Canada : Organisme statistique national du Canada. https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1710000901.

Stephenson, L. B., Harell, A., Rubenson, D., &amp; Loewen, P. J. (2020, May 1). 2019 Canadian Election Study - Online Survey. Harvard Dataverse. https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi%3A10.7910%2FDVN%2FDUS88V.

S, T. (2020, May 06). What is statistical bias and why is it so important in data science? Retrieved from https://towardsdatascience.com/what-is-statistical-bias-and-why-is-it-so-important-in-data-science-80e02bf7a88d

Urback, R. (2019, September 27). OPINION | What's the difference between the Conservative and Liberal platforms? The colour: Opinion | CBC News. CBCnews. https://www.cbc.ca/news/opinion/policy-platforms-1.5298859.

Westfall, J. A., Patterson, P. L., & Coulston, J. W. (2011). Post-stratified estimation: Within-strata and total sample size recommendations. Canadian Journal of Forest Research, 41(5), 1130-1139. doi:10.1139/x11-031


